{\bf Practical Concurrency Testing}

### 2018-10-16 thesis/defense/snakefight.tex
buncha updates from practice 2, color coding etc

\begin{document}
\renewcommand{\inserttotalframenumber}{67}
\normalem
\begin{frame}
	\titlepage
\end{frame}

@@

@@

\begin{frame}{Concurrency!}
	\begin{center}
	\begin{tabular}{ll}
		Initially \texttt{count = 0;} \\
		\\
		{\bf Thread 1} & {\bf Thread 2} \\
		\\
		\texttt{\hilight{darkorange}{mutex\_lock}(m);}      & \texttt{atomic\_xadd(\&count, 1);} \\
		\texttt{count++;}				   & \texttt{\hilight{olivegreen}{yield}();} \\
		\texttt{\hilight{darkblue}{mutex\_unlock}(m);}      & \texttt{atomic\_xadd(\&count, 1);} \\
		\texttt{assert(count >= 1);}			& \texttt{assert(count >= 2);} \\
		\\
		\\
		\\
		\\
		\\
	\end{tabular}
	\end{center}
\end{frame}
\begin{frame}{Concurrency!}
	\begin{center}
	\begin{tabular}{ll}
		Initially \texttt{count = 0;} \\
		\\
		{\bf Thread 1} & {\bf Thread 2} \\
		\\
		\texttt{\hilight{darkorange}{mutex\_lock}(m);} \\
		\texttt{int tmp = count;} \\
								& \texttt{atomic\_xadd(\&count, 1);} \\
								& \texttt{\hilight{olivegreen}{yield}();} \\
								& \texttt{atomic\_xadd(\&count, 1);} \\
		\texttt{count = tmp + 1;} \\
								& \texttt{\hilight{pinkish}{assert(count >= 2);}} \\
		\texttt{\hilight{darkblue}{mutex\_unlock}(m);} \\
		\texttt{assert(count >= 1);}
	\end{tabular}
	\end{center}
\end{frame}

\begin{frame}{Stateless Model Checking}
	Stateless Model Checking {\em [Godefroid '97]} (MC)
	\begin{itemize}
		\item Test framework controls thread scheduling
		\item Each test iteration, test a different interleaving
		\item Goal: Exhaustively check all possible program behaviours
	\end{itemize}
	\pause
	\linegap

	\begin{center}
		\includegraphics[width=0.25\textwidth]{../../oopsla/exponential-curve.pdf}
	\end{center}
\end{frame}

\begin{frame}{State Space Example}
	\begin{center}
		\includegraphics[width=0.96\textwidth]{../../oopsla/tree-maximal-only.pdf}
	\end{center}
	\linegap

	Possible interleavings represented as a tree
	\begin{itemize}
		\item Node: Intermediate execution state
		\item Edge: State transition from executing a thread
	\end{itemize}
	\linegap

@@

	Too much manual intervention %for non-expert use
	\begin{itemize}
		\item State space size estimation {\em [Simsa '12]}
			% TODO: i just switched the order of these; see how that feels
		\item Manual kernel annotations: Landslide {\em [Blum '12]} {\tiny that's me!}
			\pause
		\item Widespread adoption requires {\bf fully automated instrumentation}...
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{Thesis Statement}
	% data race pps (as well as txn equivalence)
	Combining %both
	theoretically-founded automatic reduction techniques
	% iterative deepening
	% (say smth like: "landslide also has a wealth of other user-tunable heuristics
	% for expert users, which you can read about in the document; for this talk i will focus just on the CPU budget thing")
	% XXX: need to prepare for the question -- "what evidence do you have that users actually want this cpu budget stuff",
	% ",or, that it helps vs just running something unlimited vs a normal icb test"
	% (well, for the 2nd half of the Q, i can talk about how ID outperforms ICB on verficiations)
	and user-informed heuristic ones,
	stateless model checking can sufficiently mitigate exponential explosion to be
	% education
	a practical testing technique for inexperienced users
	% htm
	% TODO: this last part you just added for this talk; does the thesis version of the statmeent need to include too?
	and real-world programs alike.
\end{frame}

\begin{frame}{Thesis Statement}
	% data race pps (as well as txn equivalence)
	\hilight{gray}{Combining %both
	theoretically-founded automatic reduction techniques
	% iterative deepening
	% (say smth like: "landslide also has a wealth of other user-tunable heuristics
	% for expert users, which you can read about in the document; for this talk i will focus just on the CPU budget thing")
	% XXX: need to prepare for the question -- "what evidence do you have that users actually want this cpu budget stuff",
	% ",or, that it helps vs just running something unlimited vs a normal icb test"
	% (well, for the 2nd half of the Q, i can talk about how ID outperforms ICB on verficiations)
	and user-informed heuristic ones, stateless model checking...}
	\begin{itemize}
		\item ...can sufficiently mitigate exponential explosion...
		% education
		\item ...to be a practical testing technique for inexperienced users...
		% htm
		% TODO: this last part you just added for this talk; does the thesis version of the statmeent need to include too?
		\item ...and real-world programs alike.
	\end{itemize}
\end{frame}

\begin{frame}{Thesis Statement}
	The Quicksand algorithm %combines theory and heuristics to
	practically debugs and verifies many exponential state spaces.
	\vspace{4.54em}

	Landslide's automatic instrumentation makes MC accessible to students.
	\vspace{4.54em}

	New reduction techniques for transactional memory keep MC up-to-date
	with the latest real-world concurrency techniques.
	\vspace{4.54em}
\end{frame}
\begin{frame}{Outline}
	\hilight{gray}{The Quicksand algorithm %combines theory and heuristics to
	practically debugs and verifies many exponential state spaces.} \\
	%\hilight{gray}{The Quicksandalgorithm makes MC appropriate for both educational and real-world programs.} \\
	\begin{itemize}
		\item {\bf Quicksand}
		\begin{itemize}
			\item Data-race preemption points \& algorithm overview
			\item Verification \& experimental results
		\end{itemize}
	\end{itemize}
	\hilight{gray}{Landslide's automatic instrumentation makes MC accessible to students.} \\
	%\hilight{gray}{Ad-hoc synchronization heuristics make MC accessible to student programmers.} \\
	\begin{itemize}
		\item {\bf Educational use}
		\begin{itemize}
			\item Pebbles \& Pintos OS projects
			\item Bug-finding, grades, \& surveys
		\end{itemize}
	\end{itemize}

	\hilight{gray}{New reduction techniques for transactional memory keep MC up-to-date
	with the latest real-world concurrency techniques.} \\
	%\hilight{gray}{A new transactional concurrency model keeps MC applicable to the latest real-world techniques.} \\
	\begin{itemize}
		\item {\bf Transactional memory}
		\begin{itemize}
			\item Concurrency model \& failure injection
			\item Bug-finding \& verification
		\end{itemize}
	\end{itemize}
\end{frame}

@@

\breakslide{\Large Quicksand}

\begin{frame}{Preemption Points}
	The burning question: ``Which {\bf preemption points} (PPs) are important?''
	\linegap

	State space of interleavings is {\em parameterized} by PPs.
	\begin{itemize}
		\item Too few PPs: Many bugs will go undetected
		\item Preempt everywhere: Completion is often infeasible % say: "for some programs"
	\end{itemize}
\end{frame}

@@

\begin{frame}{Data Race Analysis}
	A {\bf data race} is a pair of memory accesses between two threads, where:
	\begin{itemize}
		\item At least one of the accesses is a write
		\item The threads are not holding the same mutex
		\item One access {\em happens-before} the other
			%(e.g., no \texttt{cond\_signal()} in between)
			(e.g., no \texttt{cond\_wait()})
	\end{itemize}
	\pause
	\linegap

	%Prior model checkers hard-code a fixed set of PPs, committing to one state space in advance.
	Prior model checkers commit to a fixed set of PPs in advance.
	\begin{itemize}
		\item ``Preempt on {\tt mutex\_lock()}, {\tt mutex\_unlock()} only''
		\item ``Preempt on every single memory access''
	\end{itemize}
	\linegap

@@

\begin{frame}{Total Verification}
	\textbf{Convergence theorem:} Testing maximal state space, after detecting all data races, $\equiv$ preempting everywhere!
	%\begin{itemize}
	%	\item Insight: $\forall$ arbitrary buggy interleaving, $\exists$ equivalent interleaving % say 'under DPOR' i guess?
	%		using only sync/data-race PPs, \& Quicksand will eventually reach it.
	%\end{itemize}
	\linegap

@@

\begin{frame}{Summary}
	Incorporate data-race analysis to add new PPs in a feedback loop
	\linegap

	Enough time to saturate \& test every data-race PP? Full verification
	\linegap

@@

\breakslide{\Large Education}

@@

	Bug-finding
	\begin{itemize}
		\item CMU: 115 deterministic, 122 concurrency bugs in 75/103 P2s
		\item U. Chicago: 3 deterministic, 4 concurrency bugs in 3/4 kernels
			% TODO: what to do abt this bullet pts
			%\pause
		\item PSU: (no log collection due to IRB issues)
		% \item Case study tej
	\end{itemize}
\end{frame}

@@

\begin{frame}{Effect on Grades}
	\begin{center}
		\includegraphics[width=\textwidth]{photo-of-ze-studence-cdf-no-pre-study.pdf}
	\end{center}
\end{frame}

\begin{frame}{Effect on Grades}
	\begin{center}
		\includegraphics[width=\textwidth]{../photo-of-ze-studence-cdf.pdf}
	\end{center}
\end{frame}

\begin{frame}{Statistical Significance}
	Compared %users' vs. non-users'
	grade distributions with k-sample Anderson-Darling {\em [A., D. '52]}
	\linegap

	Control 1: non-users from {\bf same semester}
	\begin{itemize}
		\item {\bf P2} grade distributions differ with $p = 0.036 < 0.05$
		\item {\bf P3} grade distributions differ with $p = 0.034 < 0.05$
			\pause
		\item Possible bias: stronger students volunteer for Landslide in first place
	\end{itemize}
	\pause
	\linegap

	Control 2: non-users from {\bf past semesters} (when Landslide not offered)
	\begin{itemize}
		\item {\bf P2} grade distributions differ with $p = 0.671$...
		\item {\bf P3} grade distributions differ with $p = 0.145$...
			\pause
		\item Possible bias: grading criteria may drift across semesters
			% FIXME does this bullet belong in this sub list
		\item Possible bias: maybe Landslide finds bugs the official tests don't
	\end{itemize}
\end{frame}

@@

\begin{frame}{Survey Results (educational value)}
	\begin{center}
		\includegraphics[width=\textwidth]{../survey6.pdf} % easier to diagnose?
	\end{center}
\end{frame}

\begin{frame}{Survey Results (happiness value)}
	\begin{center}
		\includegraphics[width=\textwidth]{../survey5.pdf} % easy to understand?
	\end{center}
\end{frame}

\begin{frame}{Survey Results (educational value)}
	\begin{center}
		\includegraphics[width=\textwidth]{../survey9.pdf} % 100 cpu hours
	\end{center}
\end{frame}

\begin{frame}{Survey Results (educational value)}
	\begin{center}
		\includegraphics[width=\textwidth]{../survey8.pdf} % learned to understand concurrency
	\end{center}
\end{frame}

\begin{frame}{Survey Results (happiness value)}
	\begin{center}
		\includegraphics[width=\textwidth]{../survey10.pdf} % recommend to friend
	\end{center}
\end{frame}

@@

\breakslide{\Large Transactional Memory}

@@

	Equivalence theorem
	\begin{itemize}
		\item Safe to automatically skip all interleavings within transactions %$\qed$
			\begin{itemize}
				\item Exponential state space reduction
				\item Traditional reduction (DPOR) applied on top
			\end{itemize}
			% TODO prob just comment this out
		\item More in bonus slides...
	\end{itemize}
\end{frame}

@@

\breakslide{\Large Conclusion}

\begin{frame}{Conclusion}
	% data race pps (as well as txn equivalence)
	Combining %both
	theoretically-founded automatic reduction techniques
	% iterative deepening
	% (say smth like: "landslide also has a wealth of other user-tunable heuristics
	% for expert users, which you can read about in the document; for this talk i will focus just on the CPU budget thing")
	% XXX: need to prepare for the question -- "what evidence do you have that users actually want this cpu budget stuff",
	% ",or, that it helps vs just running something unlimited vs a normal icb test"
	% (well, for the 2nd half of the Q, i can talk about how ID outperforms ICB on verficiations)
	and user-informed heuristic ones, stateless model checking...
	\begin{itemize}
		\item ...can sufficiently mitigate exponential explosion...
		% education
		\item ...to be a practical testing technique for inexperienced users...
		% htm
		% TODO: this last part you just added for this talk; does the thesis version of the statmeent need to include too?
		\item ...and real-world programs alike.
	\end{itemize}
\end{frame}
\begin{frame}{Conclusion}
	% data race pps (as well as txn equivalence)
	Combining %both
	{\bf theoretically-founded automatic reduction techniques}
	\vspace{4.82em}

	and {\bf user-informed heuristic ones},
	\vspace{4.82em}

	stateless model checking...
	\begin{itemize}
		\item ...can sufficiently mitigate exponential explosion...
		\item ...to be a practical testing technique for inexperienced users...
		\item ...and real-world programs alike.
	\end{itemize}
\end{frame}
\begin{frame}{Conclusion}
	\hilight{gray}{Combining %both
	{\bf theoretically-founded automatic reduction techniques}}
	\begin{itemize}
		\item Verification with data-race preemption points
		\item Student yield-loop blocking % student (adj). from "studence"
		\item Equivalent emulation of HTM semantics
	\end{itemize}
	\hilight{gray}{and {\bf user-informed heuristic ones},}
	\begin{itemize}
		\item Quicksand's CPU budget argument
		\item Exposing data-race reports to students
		\item Abstraction reduction for HTM primitives
	\end{itemize}
	\hilight{gray}{stateless model checking...}
	\begin{itemize}
		\item \hilight{gray}{...can sufficiently mitigate exponential explosion...}
		\item \hilight{gray}{...to be a practical testing technique for inexperienced users...}
		\item \hilight{gray}{...and real-world programs alike.}
	\end{itemize}
\end{frame}
\begin{frame}{Closing remarks}
	``Verifying C programs is doomed, why can't PL design handle this?''
	\begin{itemize}
		\item seL4 {\em [Klein '09]}: secure kernel verified end-to-end in Haskell
		\item Rust {\em [Matsakis '14]}: concurrency-aware type system, no data races
	\end{itemize}
	\pause
	\linegap

@@

\subsection{Quicksand}
\breakslide{\Large Bonus Slides - Quicksand}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Education}
\breakslide{\Large Bonus Slides - Education}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{HTM}
\breakslide{\Large Bonus Slides - HTM}

>>>>

\definecolor{sect-quicksand}{RGB}{0,96,128}
\definecolor{sect-410}{RGB}{0,128,64}
\definecolor{sect-htm}{RGB}{128,0,64}
\definecolor{sect-pastel-quicksand}{RGB}{144,180,192}
\definecolor{sect-pastel-410}{RGB}{144,192,168}
\definecolor{sect-pastel-htm}{RGB}{192,144,168}

\begin{document}
\renewcommand{\inserttotalframenumber}{69}
\normalem
\begin{frame}
	\titlepage
\end{frame}

@@

% TODO: on the title slide, shouldn't there be a "PResented in partial requirement..."?

@@

\begin{frame}{Concurrency}
	\begin{center}
	\begin{tabular}{ll}
		Initially \texttt{count = 0;} \\
		\\
		{\bf Thread 1} & {\bf Thread 2} \\
		\\
		\texttt{\hilight{darkorange}{mutex\_lock}(m);}      & \texttt{atomic\_xadd(\&count, 1);} \\
		\texttt{count++;}				   & \texttt{\hilight{olivegreen}{yield}();} \\
		\texttt{\hilight{darkblue}{mutex\_unlock}(m);}      & \texttt{atomic\_xadd(\&count, 1);} \\
		\texttt{assert(count >= 1);}			& \texttt{assert(count >= 2);} \\
		\\
		\\
		\\
		\\
		\\
	\end{tabular}
	\end{center}
\end{frame}
\begin{frame}{Concurrency}
	\begin{center}
	\begin{tabular}{ll}
		Initially \texttt{count = 0;} \\
		\\
		{\bf Thread 1} & {\bf Thread 2} \\
		\\
		\texttt{\hilight{darkorange}{mutex\_lock}(m);} \\
		\texttt{int tmp = count;} \\
								& \texttt{atomic\_xadd(\&count, 1);} \\
								& \texttt{\hilight{olivegreen}{yield}();} \\
								& \texttt{atomic\_xadd(\&count, 1);} \\
		\texttt{count = tmp + 1;} \\
								& \texttt{\hilight{pinkish}{assert(count >= 2);}} \\
		\texttt{\hilight{darkblue}{mutex\_unlock}(m);} \\
		\texttt{assert(count >= 1);}
	\end{tabular}
	\end{center}
\end{frame}

\begin{frame}{Concurrency Testing} % "..is hard"?
	% what's so bad abt concurrecny bugs
	% - manifest at random depending on interrupts/etc
	%   - (fine in testing but in production, loaded server, more interrupts, more interleavings, more crashes)
	% - difficult to reproduce
	% - difficult to know yr testing "coverage" - did you *really* fix that bug, or just make it more rare?
	% - difficult to use, as evidenced by most systems classes still teaching kids to use stress (or worse, unit) tests
	Concurrency bugs...
	\begin{itemize}
		\item Manifest at random
		\item Difficult to reproduce
		\item Difficult to know when truly fixed % what is ``coverage''?
	\end{itemize}
	\linegap

	Testing approaches
	\begin{itemize}
		\item Unit tests: only 1 ``most likely'' interleaving
		\item Stress tests: exposes ``less likely'' interleavings (if lucky), % to expose "less likely" ones
			cannot verify
		\item Verification: steep learning curve, unpredictable runtime
			%still no widespread adoption % -- why?
	\end{itemize}
\end{frame}
\begin{frame}{Practical Concurrency Testing}
	%Goals
	\textbf{Landslide}, this thesis's solution
	\begin{itemize}
		\item No randomness, despite nondeterminism
		\item Confirm when you fixed bugs, not just find them % not 'verify'
			% TODO: make alnguage match, "Automatic so and so"
		\item User-friendly enough for students
		% \item Test wide variety of types of concurrency % txn
			% TODO : figure out how to ref 
		% ????
		\item Combines with human intuition to verify tests neither could do alone
	\end{itemize}
	%
	% enter landslide
	% - test yr nondeterministic program without randomness/unpredictability
	% - provide formal verification guarantees so you know when you actually fixed that bug
	%   - (well, subject to assumptions about the test case you're running under, etc etc)
	% - user friendly enough even for ze studence
	% 
	% say: "What if there was a program that... ...I'm here to tell you that I've built such a program,
	% it's called Landslide, and why you should let me graduate for it" (figure out smth more diplomatic here...)
\end{frame}

\begin{frame}{Stateless Model Checking}
	Stateless Model Checking (MC) {\em [Godefroid '97]}
	\begin{itemize}
		\item Test framework controls thread scheduling
		\item Each test iteration, test a different interleaving
		\item Goal: Exhaustively check all possible program behaviours
	\end{itemize}
	\pause
	\linegap

	\begin{center}
		\includegraphics[width=0.4\textwidth]{exponence.pdf}
	\end{center}
\end{frame}

% better: remember to say, "If you're wondering what the deal is with "stateless"...
% the name is a historical accident & we can talk abt that in the Q+As"
\begin{frame}{Stateless Model Checking - The State Space}
	\begin{center}
		\includegraphics[width=0.96\textwidth]{../../oopsla/tree-maximal-only.pdf}
	\end{center}
	\linegap

	Possible interleavings (under given test case) represented as a tree
	\begin{itemize}
		\item Node: Intermediate execution state
		\item Edge: State transition from executing a thread
	\end{itemize}
	\linegap

@@

	Too much manual intervention %for non-expert use
	\begin{itemize}
		\item State space size estimation {\em [Simsa '12]}
			% TODO: i just switched the order of these; see how that feels
		\item Manual kernel annotations {\em [Blum '12]} %{\tiny that's me!}
			\pause
		\item Widespread adoption requires {\bf fully automated instrumentation}...
	\end{itemize}
	\pause
	\linegap

@@

\begin{frame}{Thesis Statement}
	% data race pps (as well as txn equivalence)
	Combining %both
	theoretically-founded automatic reduction techniques
	% iterative deepening
	% (say smth like: "landslide also has a wealth of other user-tunable heuristics
	% for expert users, which you can read about in the document; for this talk i will focus just on the CPU budget thing")
	% XXX: need to prepare for the question -- "what evidence do you have that users actually want this cpu budget stuff",
	% ",or, that it helps vs just running something unlimited vs a normal icb test"
	% (well, for the 2nd half of the Q, i can talk about how ID outperforms ICB on verficiations)
	and user-informed heuristic ones,
	stateless model checking can sufficiently mitigate exponential explosion to be
	% education
	practical for inexperienced users
	%a practical testing technique for inexperienced users
	% htm
	% TODO: this last part you just added for this talk; does the thesis version of the statmeent need to include too?
	and real-world programs alike.
\end{frame}

\begin{frame}{Thesis Statement}
	% data race pps (as well as txn equivalence)
	\hilight{gray}{Combining %both
	theoretically-founded automatic reduction techniques
	% iterative deepening
	% (say smth like: "landslide also has a wealth of other user-tunable heuristics
	% for expert users, which you can read about in the document; for this talk i will focus just on the CPU budget thing")
	% XXX: need to prepare for the question -- "what evidence do you have that users actually want this cpu budget stuff",
	% ",or, that it helps vs just running something unlimited vs a normal icb test"
	% (well, for the 2nd half of the Q, i can talk about how ID outperforms ICB on verficiations)
	and user-informed heuristic ones, stateless model checking...}
	\begin{itemize}
		\item \hilight{sect-quicksand}{...can sufficiently mitigate exponential explosion...}\xspace
		% education
		%\item ...to be a practical testing technique for inexperienced users...
		\item \hilight{sect-410}{...to be practical for inexperienced users...}\xspace
		% htm
		% TODO: this last part you just added for this talk; does the thesis version of the statmeent need to include too?
		\item \hilight{sect-htm}{...and real-world programs alike.}\xspace
	\end{itemize}
\end{frame}

\begin{frame}{Contributions}
	\hilight{sect-quicksand}{The Quicksand algorithm %combines theory and heuristics to
	practically debugs and verifies many exponential state spaces.}
	\vspace{4.54em}

	\hilight{sect-410}{Landslide's automatic instrumentation makes MC accessible to students.}
	\vspace{4.54em}

	\hilight{sect-htm}{New reduction techniques for transactional memory keep MC up-to-date
	with the latest real-world concurrency techniques.}
	\vspace{4.54em}
\end{frame}
\begin{frame}{Outline}
	\hilight{sect-pastel-quicksand}{The Quicksand algorithm %combines theory and heuristics to
	practically debugs and verifies many exponential state spaces.} \\
	%\hilight{gray}{The Quicksandalgorithm makes MC appropriate for both educational and real-world programs.} \\
	\begin{itemize}
		\item {\bf Quicksand}
		\begin{itemize}
			\item Data-race preemption points \& algorithm overview
			\item Verification \& experimental results
		\end{itemize}
	\end{itemize}
	\hilight{sect-pastel-410}{Landslide's automatic instrumentation makes MC accessible to students.} \\
	%\hilight{gray}{Ad-hoc synchronization heuristics make MC accessible to student programmers.} \\
	\begin{itemize}
		\item {\bf Educational use}
		\begin{itemize}
			\item Pebbles \& Pintos OS projects
			\item Bug-finding, grades, \& surveys
		\end{itemize}
	\end{itemize}

	\hilight{sect-pastel-htm}{New reduction techniques for transactional memory keep MC up-to-date
	with the latest real-world concurrency techniques.} \\
	%\hilight{gray}{A new transactional concurrency model keeps MC applicable to the latest real-world techniques.} \\
	\begin{itemize}
		\item {\bf Transactional memory}
		\begin{itemize}
			\item Concurrency model \& failure injection
			\item Bug-finding \& verification
		\end{itemize}
	\end{itemize}
\end{frame}

@@

\breakslide{\Large \bf \hilight{sect-quicksand}{Quicksand}}

\begin{frame}{Preemption Points}
	The burning question: ``Which {\bf preemption points} are important?''
	\linegap

	State space %of interleavings
	is {\em parameterized} by preemption points.
	\begin{itemize}
		\item Too few: Many bugs will go undetected
		\item Preempt everywhere: Completion is often infeasible % say: "for some programs"
	\end{itemize}
\end{frame}

@@

\begin{frame}{Data Race Analysis}
	A {\bf data race} is a pair of memory accesses between two threads, where:
	\begin{itemize}
		\item At least one of the accesses is a write
		\item The threads are not holding the same mutex
		\item Neither access {\em happens-before} the other
			%(e.g., no \texttt{cond\_signal()} in between)
			(e.g., no \texttt{cond\_wait()})
	\end{itemize}
	\pause
	\linegap

	%Prior model checkers hard-code a fixed set of PPs, committing to one state space in advance.
	%Prior model checkers commit to a fixed set of PPs in advance.
	Prior model checkers commit to fixed preemption points in advance.
	\begin{itemize}
		\item ``Preempt on {\tt mutex\_lock()}, {\tt mutex\_unlock()} only''
		\item ``Preempt on every single memory access''
	\end{itemize}
	\linegap

@@

\begin{frame}{Total Verification}
	\textbf{Convergence theorem:} Testing maximal state space, after detecting all data races, $\equiv$ preempting everywhere!
	\begin{itemize}
		\item $\forall$ arbitrary buggy interleaving, $\exists$ equivalent interleaving % say 'under DPOR' i guess?
			using only sync/data-race PPs, \& Quicksand will eventually reach it.
			\begin{itemize}
				\item ({\em equivalence} defined under DPOR {\em [Flanagan '05]})
			\end{itemize}
	\end{itemize}
	\linegap

@@

\begin{frame}{Summary}
	Use data-race analysis to add new preemption points in a feedback loop
	\linegap

	Enough time to saturate \& test every data-race? Full verification
	\linegap

@@

\breakslide{\Large \bf \hilight{sect-410}{Education}}

@@

	Bug-finding
	\begin{itemize}
		\item CMU: 115 deterministic, 122 concurrency bugs in 75/103 P2s
		\item U. Chicago: 3 deterministic, 4 concurrency bugs in 3/4 kernels
			% TODO: what to do abt this bullet pts
			%\pause
		\item PSU: (no log collection) %due to IRB issues)
		% \item Case study tej
	\end{itemize}
\end{frame}

@@

% TODO: put p values
\begin{frame}{Effect on Grades (CMU)} % - statistically significant?}
	\begin{center}
		\includegraphics[width=\textwidth]{photo-of-ze-studence-cdf-no-pre-study.pdf}
	\end{center}
\end{frame}

\begin{frame}{Effect on Grades (CMU)} % - statistically significant?}
	\begin{center}
		\includegraphics[width=\textwidth]{../photo-of-ze-studence-cdf.pdf}
	\end{center}
\end{frame}

%\begin{frame}{Statistical Significance}
%	Compared %users' vs. non-users'
%	grade distributions with k-sample Anderson-Darling {\em [A., D. '52]}
%	\linegap
%
%	Control 1: non-users from {\bf same semester}
%	\begin{itemize}
%		\item {\bf P2} grade distributions differ with $p = 0.036 < 0.05$
%		\item {\bf P3} grade distributions differ with $p = 0.034 < 0.05$
%			\pause
%		\item Possible bias: stronger students volunteer for Landslide in first place
%	\end{itemize}
%	\pause
%	\linegap
%
%	Control 2: non-users from {\bf past semesters} (when Landslide not offered)
%	\begin{itemize}
%		\item {\bf P2} grade distributions differ with $p = 0.671$...
%		\item {\bf P3} grade distributions differ with $p = 0.145$...
%			\pause
%		\item Possible bias: grading criteria may drift across semesters
%			% FIXME does this bullet belong in this sub list
%		\item Possible bias: maybe Landslide finds bugs the official tests don't
%	\end{itemize}
%\end{frame}

@@

\begin{frame}{Survey Results} % (educational value)}
	\begin{center}
		\includegraphics[width=\textwidth]{../survey6.pdf} % easier to diagnose?
	\end{center}
\end{frame}

%\begin{frame}{Survey Results} % (happiness value)}
%	\begin{center}
%		\includegraphics[width=\textwidth]{../survey5.pdf} % easy to understand?
%	\end{center}
%\end{frame}

\begin{frame}{Survey Results} % (educational value)}
	\begin{center}
		\includegraphics[width=\textwidth]{../survey9.pdf} % 100 cpu hours
	\end{center}
\end{frame}

\begin{frame}{Survey Results} % (educational value)}
	\begin{center}
		\includegraphics[width=\textwidth]{../survey8.pdf} % learned to understand concurrency
	\end{center}
\end{frame}

\begin{frame}{Survey Results} % (happiness value)}
	\begin{center}
		\includegraphics[width=\textwidth]{../survey10.pdf} % recommend to friend
	\end{center}
\end{frame}

@@

\breakslide{\Large \bf \hilight{sect-htm}{Transactional Memory}}

@@

	Equivalence theorem
	\begin{itemize}
		\item Safe to automatically skip all interleavings within transactions %$\qed$
			\begin{itemize}
				\item Exponential state space reduction
				\item Traditional reduction (DPOR) applied on top
			\end{itemize}
		%\item More in bonus slides...
	\end{itemize}
\end{frame}

@@

\breakslide{\Large \bf Conclusion}

\begin{frame}{Conclusion}
	% data race pps (as well as txn equivalence)
	Combining %both
	theoretically-founded automatic reduction techniques
	% iterative deepening
	% (say smth like: "landslide also has a wealth of other user-tunable heuristics
	% for expert users, which you can read about in the document; for this talk i will focus just on the CPU budget thing")
	% XXX: need to prepare for the question -- "what evidence do you have that users actually want this cpu budget stuff",
	% ",or, that it helps vs just running something unlimited vs a normal icb test"
	% (well, for the 2nd half of the Q, i can talk about how ID outperforms ICB on verficiations)
	and user-informed heuristic ones,
	\linegap

	stateless model checking
	\hilight{sect-quicksand}{can sufficiently mitigate exponential explosion}\xspace
	\hilight{sect-410}{to be practical for inexperienced users}\xspace
	\hilight{sect-htm}{and real-world programs alike.}\xspace
\end{frame}
\begin{frame}{Conclusion}
	% data race pps (as well as txn equivalence)
	Combining %both
	{\bf theoretically-founded automatic reduction techniques}
	\vspace{4.82em}

	and {\bf user-informed heuristic ones},
	\vspace{4.82em}
	\linegap

	stateless model checking
	\hilight{sect-quicksand}{can sufficiently mitigate exponential explosion}\xspace
	\hilight{sect-410}{to be practical for inexperienced users}\xspace
	\hilight{sect-htm}{and real-world programs alike.}\xspace
\end{frame}
\begin{frame}{Conclusion}
	%\hilight{gray}{Combining %both
	%{\bf theoretically-founded automatic reduction techniques}}
	Combining %both
	{\bf theoretically-founded automatic reduction techniques}
	\begin{itemize}
		\item \hilight{sect-quicksand}{Verification with data-race preemption points}
		\item \hilight{sect-410}{Student yield-loop blocking} % student (adj). from "studence"
		\item \hilight{sect-htm}{Equivalent emulation of HTM semantics}
	\end{itemize}
	% TODO: incremental reveal
	%\hilight{gray}{and {\bf user-informed heuristic ones},}
	and {\bf user-informed heuristic ones},
	\begin{itemize}
		\item \hilight{sect-quicksand}{Quicksand's CPU budget argument}
		\item \hilight{sect-410}{Exposing data-race reports to students}
		%\item Encouraging students to exercise their intuition % ??
		\item \hilight{sect-htm}{Abstraction reduction for HTM primitives}
	\end{itemize}
	\linegap

	\hilight{gray}{stateless model checking}\xspace
	\hilight{sect-pastel-quicksand}{can sufficiently mitigate exponential explosion}\xspace
	\hilight{sect-pastel-410}{to be practical for inexperienced users}\xspace
	\hilight{sect-pastel-htm}{and real-world programs alike.}\xspace
\end{frame}
\begin{frame}{Closing remarks}
	``Verifying C programs is doomed, why can't PL design handle this?''
	\begin{itemize}
		\item seL4 {\em [Klein '09]}: secure kernel verified end-to-end in Haskell
		\item Rust {\em [Matsakis '14]}: concurrency-aware type system, no data races
	\end{itemize}
	\pause
	\linegap

@@

\subsection{Quicksand}
\breakslide{\Large \bf \hilight{sect-quicksand}{Bonus Slides - Quicksand}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Education}
\breakslide{\Large \bf \hilight{sect-410}{Bonus Slides - Education}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{HTM}
\breakslide{\Large \bf \hilight{sect-htm}{Bonus Slides - HTM}}
